---
title: "R Notebook to illustrate using R packages to predict race for voters without geocoding"
output: html_notebook
---

```{r}
library(tidyverse)
library(rethnicity)
library(predictrace)
library(zipWRUext2)
```

load the Queens voter registration records came in fixed width format
```{r}
width = c(9,30,30,1,4,10,10,15,50,
          40,5,4,50,50,50,50,8,
          1,3,30,3,2,2,2,2,2,2,
          8,2,1,8,4,20,3,15,8,100)
df = read.fwf("Queens.txt", widths = width, fill = TRUE)
str(df) #1190288 obs. of  37 variables

# add column names and replace the space with "_"
colnames(df) = c("County EMSID", "Last Name", "First Name", "Middle Initial","Name Suffix",
                 "House Number","House Number Suffix","Apartment Number","Street Name","City","Zip Code",
                 "Zip Code4","Mailing Address 1","Mailing Address 2","Mailing Address 3",
                 "Mailing Address 4","Birth Date","Gender", "Political Party", "Other Party",
                 "Election District","Assembly District","Congress District","Council District",
                 "Senate District","Civil Court District","Judicial District","Registration Date",
                 "Status Code","Voter Type","Eff Status Change Date","Year Last Voted",
                 "TelephoneNum","Future Party","Future Other Party",
                 "Future Party Effective Date","Email Address")

colnames(df) = gsub(" ", "_", colnames(df))

df_copy = df
                 
str(df_copy) #1190288 obs. of  37 variables
```


check missing values
```{r}
missing_counts <- df_copy %>%
  summarise(across(everything(), ~sum(is.na(.))))

print(missing_counts)

# check if any duplicates in column County_EMSID
duplicates <- sum(duplicated(df$County_EMSID))
print(duplicates)

```


check how the prediction packages work
predict_ethnicity method only accepts fullname or lastname
```{r}
test = predict_ethnicity(lastnames = "YU", method="lastname")
test

test1 = predictrace::predict_race("Yu", probability = FALSE)[,3]
test1

test2 = predictrace::predict_race("Yue", surname = FALSE, probability = FALSE)[,3]
test2

```


Create race1 and race2 columns using rethnicity and predictrace packages
```{r}
df_copy <- df_copy %>%
  mutate(race1 = predict_ethnicity(lastnames = Last_Name, method="lastname")[,6],
         race2 = predictrace::predict_race(Last_Name, probability = FALSE)[,3]
)

head(df_copy, 100)

```


Create race3 column using zipWRUext2 package
```{r}
df_copy = zip_wru(df_copy, state = "NEW YORK",
                             year1 = "2020", zip_col = "Zip_Code", surname_field = "Last_Name")

# below section is to find the race that has the maximum probability
# Specify the columns for which you want to find the maximum
cols_to_check <- c("pred.whi", "pred.bla", "pred.his", "pred.asi", "pred.oth")

# Get the column index of the maximum value in each row across specified columns
max_col_index <- max.col(df_copy[cols_to_check])
max_col_index

# Create a new column to flag which column contains the maximum value
df_copy$max_prob_race <- cols_to_check[max_col_index]

# Print the data frame with the max_column flag
#print(stratified_df_test)

#class(stratified_df_test$max_prob_race)

# create column race3
df_copy = df_copy %>%
  mutate(race3 = case_when(
    str_detect(max_prob_race, "\\.whi$") ~ "white",
    str_detect(max_prob_race, "\\.asi$") ~ "asian",
    str_detect(max_prob_race, "\\.his$") ~ "hispanic",
    str_detect(max_prob_race, "\\.bla$") ~ "black",
    str_detect(max_prob_race, "\\.oth$") ~ "other",
  ))

head(df_copy, 10)

# replace NA in race3 column with string "other"
df_copy$race3 = ifelse(is.na(df_copy$race3), "other", df_copy$race3)
df_copy %>% count(race3, name="UniqueNames")
#df_copy %>% count(party_full, name = "UniqueCount")
```

compare the prediction results among three race columns
```{r}
df_copy %>% count(race1, name = "UniqueCount") # does not have "other" category which doesn't make much sense
df_copy %>% count(race2, name = "UniqueCount") # too many categories
df_copy %>% count(race3, name = "UniqueCount") # probably a good idea to go with the predictions made by race3

```

Create bar chart to show the race proportion in Queens (determined by race3 column)
```{r}
df_copy %>%
  group_by(race3) %>%
  summarise(cnt = n()) %>%
  mutate(proportion = (cnt/sum(cnt))*100) %>%
  ggplot(aes(y=proportion, x=fct_reorder(race3, proportion, .desc=TRUE))) +
  geom_bar(position="dodge", stat="identity", fill="skyblue") +
  geom_text(aes(label = paste0(round(proportion,1), "%")), vjust = -0.5, color = "black", size = 3) +
  labs(title = "Race % in Queens County New York",x = "race") +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(color = "black"),
        axis.text.x = element_text(angle = 45, hjust = 1))
```
