---
title: "R Notebook for Queens county voting history data cleanup"
output: html_notebook
---

```{r}
library(tidyverse)
library(ggpattern)
library(hrbrthemes)
library(viridis)
library(janitor)
```




load the voting history file came in fixed width format
```{r}
width = c(9,1,2,3,3,8,2,1,1)

df_voter_hisotry = read.fwf("Queens History 11072023.txt", widths = width, fill = TRUE)

# add column names
colnames(df_voter_hisotry) = c("county_emsid", "Boro", "Assembly_District", "Election_District","Political_Party",
                 "Election_Date","Election_Type","Voter_Type","Voting_Method")


str(df_voter_hisotry)
View(df_voter_hisotry)

df_voter_hisotry_copy = df_voter_hisotry
#df_voter_hisotry_copy$County_EMSID = factor(df_voter_hisotry_copy$County_EMSID)
str(df_voter_hisotry_copy) #9913179
View(df_voter_hisotry_copy)

# below two dataframes have different memory so updating one won't update the other
tracemem(df_voter_hisotry)==tracemem(df_voter_hisotry_copy)
```

Data wrangling - part I 
1. Remove white space in county_emsid column
2. Check if the dataset has duplicated rows and remove duplicated rows
3. Extract year from Election_Date column
```{r}
df_voter_hisotry_copy = df_voter_hisotry_copy %>%
  mutate_at(vars(county_emsid), trimws)

# check duplicates - 16 rows 
#get_dupes(df_voter_hisotry_copy) # from janitor package

# two types of duplicated rows: 1) duplicate on all the fields 2) voter has two records for the same election type and election year. the only difference is Voting_Method B vs R. Type 1) can be dealt with using distinct()

#example of duplicated rows before removing
# type1
# df_voter_hisotry_copy %>%
#   filter(county_emsid=="01181591" & Election_Date==20021105)
# # type2 votes double counting
# df_voter_hisotry_copy %>%
#   filter(county_emsid=="300652999" & Election_Date==19921103)

# remove duplicate on all the fields
df_voter_hisotry_copy = df_voter_hisotry_copy %>%
  distinct()

df_voter_hisotry_copy #9,913,171 observations

# to deal with votes double counting issue caused by Voting_Method column. I decided to drop the column and then utilize distinct() one more time

df_voter_hisotry_copy = df_voter_hisotry_copy %>%
  select(-Voting_Method, -Voter_Type, -Political_Party) %>%
  #get_dupes() #18 rows
  distinct() %>%
  mutate(election_year = as.integer(substr(as.character(Election_Date), 1, 4)))

df_voter_hisotry_copy #9,913,162 observations

```

checking missingness of the columns
```{r}
df_voter_hisotry_copy %>%
  summarise(across(everything(), ~sum(is.na(.)))) #gender & race_est have 14867 missing values
```

take a quick look at the data. looks like the data includes voting history outside of Queens county
```{r}
df_voter_hisotry_copy %>% count(Boro, name = "UniqueCount")
df_voter_hisotry_copy %>% count(Election_Type, name = "UniqueCount")
#View(df_voter_hisotry_copy)
```

In real life, it is probably a good idea to exact the voting records for all the voting years. The way NGP VAN's VoteBuilder platform has in the voting history filter is that they all users to choose Any, All, At least, Exactly, No more than X times within a certain period for a certain type of election. For this project, I did 2020 to 2023 based on Wendy Li's campaign ask.

calculate how many times a voter voted in the primary elections from year 2020 to year 2023. I chose year 2020 to year 2023 b/c it convered an election cycle (year 2020 was the last year that the Presidential Election was held
)
```{r}
count_primary = as.data.frame(df_voter_hisotry_copy %>%
  filter((Election_Type %in% c("PP", "PR")) & election_year>2019) %>%
  select(county_emsid, election_year) %>% #606,983 observations
  group_by(county_emsid, election_year) %>%
  count() %>%
  pivot_wider(names_from = "election_year", names_prefix = "primary_", values_from = "n")
)

# replace NA with value zero
count_primary[is.na(count_primary)] = 0

# arrange the order of columns
count_primary = count_primary[, c("county_emsid", "primary_2020",
                                  "primary_2021", "primary_2022", "primary_2023")]
# check duplicates
count_primary %>%
  get_dupes() #no dupliates

# cross check sum matches sum before the pivot_wider operation
colSums(count_primary[c("primary_2020","primary_2022","primary_2021","primary_2023")])
sum(colSums(count_primary[c("primary_2020","primary_2022","primary_2021","primary_2023")])) #606983

count_primary #310,307 observations
```

Campaign ask: list years that voters voted from 2021 to 2023 (primary elections only)
```{r}
count_primary %>%
  filter(!(primary_2021==0 & primary_2022==0 & primary_2023==0)) %>%
  select(county_emsid, primary_2021, primary_2022, primary_2023) %>%
  write.csv(file = "voting_history_primarycount21to23.csv", row.names = FALSE)

```


calculate how many times a voter voted in the general elections from year 2020 to year 2023
```{r}
count_general = as.data.frame(df_voter_hisotry_copy %>%
  filter(Election_Type=="GE" & election_year>2019) %>%
  select(county_emsid, election_year) %>% #1,576,830 observations
  group_by(county_emsid, election_year) %>%
  count() %>%
  pivot_wider(names_from = "election_year", names_prefix = "general_", values_from = "n")
)

# replace NA with value zero
count_general[is.na(count_general)] = 0

# arrange the order of columns
count_general = count_general[, c("county_emsid", "general_2020",
                                  "general_2021", "general_2022", "general_2023")]
# check duplicates
count_general %>%
  get_dupes() #no dupliates

# cross check sum matches sum before the pivot_wider operation
colSums(count_general[c("general_2020","general_2021", "general_2022", "general_2023")])
sum(colSums(count_general[c("general_2020","general_2021", "general_2022", "general_2023")])) #1576830

count_general #767,109 observations
```

full join count_general and count_primary
```{r}
count_primary %>%
  mutate(is_in_general = county_emsid %in% count_general$county_emsid) %>%
  filter(is_in_general==FALSE)

count_join = full_join(count_general, count_primary, by = "county_emsid")

# replace NA with value zero
count_join[is.na(count_join)] = 0

count_join %>%
  get_dupes() # no duplicates which is expected

count_join #777,897 observations


```

calculate how many times a voter voted from year 2020
```{r}
count_table_since2020 = as.data.frame(df_voter_hisotry_copy %>%
                                        filter(election_year>2019) %>%
                                        select(county_emsid, Election_Type) %>%
                                        group_by(county_emsid, Election_Type) %>% #9,913,162
                                        count() %>%
                                        pivot_wider(names_from = "Election_Type", values_from = "n")
  
)

count_table_since2020

# replace NA with zero 
count_table_since2020[is.na(count_table_since2020)] = 0
#names(count_table_since2020)

# calculate below new columns
count_table_since2020 = count_table_since2020 %>%
  mutate(Last4yr_times_voted = GE+PP+PR+SP) %>%
  select(county_emsid, Last4yr_times_voted)

count_table_since2020

  
```



calculate how many times a voter voted since the inception of the voting history records
```{r}
count_table = as.data.frame(df_voter_hisotry_copy %>%
                              select(county_emsid, Election_Type) %>%
                              group_by(county_emsid, Election_Type) %>% #9,913,162
                              count() %>%
                              pivot_wider(names_from = "Election_Type", values_from = "n")
  
)

# replace NA with zero 
count_table[is.na(count_table)] = 0

# calculate below new columns
count_table = count_table %>%
  mutate(Total_times_voted = GE+PP+PR+SP+RO+SB,
         Total_times_voted_Primary = PP+PR,
         Total_times_voted_General = GE)

# double check if the total number of times voted matches the original dataset
sum(count_table$Total_times_voted) #9913162
count_table
  
```

create final extract to be merged with voter registration records
```{r}
# all observations in count_join should be in count_table but just double checking
count_join %>%
  mutate(is_in_table = county_emsid %in% count_table$county_emsid) %>%
  filter(is_in_table==FALSE) # return zero records, yay

#left join count_table with count_join and count_table_since2020
count_table = count_table %>%
  left_join(count_join, by="county_emsid") %>%
  mutate(Last4yr_times_voted_primary = primary_2020+primary_2021+primary_2022+primary_2023,
         Last4yr_times_voted_general = general_2020+general_2021+general_2022+general_2023) %>%
  left_join(count_table_since2020, by="county_emsid")

# replace NA with zero 
count_table[is.na(count_table)] = 0
count_table

# create extract
count_table %>%
  select(county_emsid, Total_times_voted, Total_times_voted_Primary,
         Total_times_voted_General, Last4yr_times_voted_primary,
         Last4yr_times_voted_general, Last4yr_times_voted) %>%
  write.csv(file = "voting_history_processed.csv", row.names = FALSE)

```


Visualizations:
note the difference between calculating how many people voted vs how many times each person voted
calculate how many people voted in general election per year

load the voters_gender_race extract from the voter registeration records
```{r}
df_gender_race = read_csv("voters_gender_race.csv")
df_gender_race

```

Data cleanup
1. Check if any county_emsid in the voting history dataset are not in the voter registeration dataset
3. Pull gender and race columns from the voter registeration records
```{r}
df_voter_hisotry_copy = df_voter_hisotry_copy %>%
  mutate(is_in_reg = county_emsid %in% df_gender_race$county_emsid) %>%
  left_join(df_gender_race, by = "county_emsid")


df_voter_hisotry_copy %>%
  get_dupes()

```

checking missingness of the columns
```{r}
df_voter_hisotry_copy %>%
  summarise(across(everything(), ~sum(is.na(.)))) #gender & race_est have 14867 missing values

df_voter_hisotry_copy %>%
  count(is_in_reg, name="UniqueNames") #14867 FALSE
```


```{r}
df_time_series_GE = df_voter_hisotry_copy %>%
  filter(Election_Type=="GE") %>%
  select(county_emsid, gender, race_est, election_year) %>%
  distinct() # to remove duplicates

# check again if data frame has duplicated rows
# one way to check it
print(df_time_series_GE[duplicated(df_time_series_GE) | duplicated(df_time_series_GE, fromLast = TRUE), ])

# another way to check it
df_time_series_GE %>%
  get_dupes()

```

create viz to show number of voters voted in general elections by gender
More females voted than males
```{r}
df_time_series_GE %>%
  filter(gender %in% c("F", "M")) %>%
  select(election_year, gender) %>%
  group_by(election_year, gender) %>%
  mutate(cnt = n()) %>%
  ggplot(aes(x=election_year, y=cnt, group=gender, color=gender)) +
  geom_line() +
    scale_color_viridis(discrete = TRUE) +
    ggtitle("Number of voters voted in general elections by gender") +
    theme_classic() +
    xlab("election year") +
    ylab("total count")
```

create viz to show general trend. total number of people voted peaked every four years on the year where Presidential election was held
data points refer to all the years where we had presidential elections
```{r}
df_presidential = df_time_series_GE %>%
  filter(election_year%% 4 ==0) %>%
  select(election_year) %>%
  group_by(election_year) %>%
  summarise(cnt = n())


df_midterm = df_time_series_GE %>%
  filter(election_year %% 2 ==0 & election_year %% 4 !=0) %>%
  select(election_year) %>%
  group_by(election_year) %>%
  summarise(cnt = n())

df_time_series_GE %>%
  filter(election_year>1984) %>%
  select(election_year) %>%
  group_by(election_year) %>%
  summarise(cnt = n()) %>%
  ggplot(aes(election_year, cnt)) + 
  geom_line(color="grey") +
  geom_point(show.legend = TRUE, data=df_presidential, shape=20, color="darkblue", size=3) +
  geom_point(show.legend = TRUE, data=df_midterm, shape=20, color="darkblue", size=2) +
  theme_classic() +
  ggtitle("Number of voters voted in the general elections each year") +
  labs(x = "year", y = "number of voters voted")

ggsave("viz/total people voted in general elections.png")
```

create viz to show number of voters voted in general elections by race
```{r}
df_time_series_GE %>%
  filter(!is.na(race_est) & election_year>1990) %>%
  select(election_year, race_est) %>%
  group_by(election_year, race_est) %>%
  mutate(cnt = n()) %>%
  ggplot(aes(x=election_year, y=cnt, group=race_est, color=race_est)) +
  geom_line() +
    scale_color_viridis(discrete = TRUE, option="H") +
    ggtitle("Number of voters voted in general elections by race") +
    theme_classic() +
    xlab("race") +
    ylab("total count")
```

personally I don't think area map is a good choice here. Line graphs (see above) does a better job
```{r}
df_time_series_GE %>%
  filter(!is.na(race_est) & election_year>1990) %>%
  select(election_year, race_est) %>%
  group_by(election_year, race_est) %>%
  mutate(cnt = n()) %>%
  ggplot(aes(x=election_year, y=cnt, fill=race_est)) +
    geom_area(alpha=0.6 , size=.5, colour="white") +
    scale_fill_viridis(discrete = T) +
    theme_ipsum() + 
    ggtitle("Number of voters voted in general elections by race") +
    xlab("election year") +
    ylab("total count")


```


calculate how many people voted in primary election per year
not much insight can be drawn from below plot
```{r}

df_PP = df_voter_hisotry_copy %>%
  filter(Election_Type %in% c("PP", "PR") & election_year%% 4 ==0) %>%
  select(county_emsid, gender, race_est, election_year) %>%
  distinct() %>%
  group_by(election_year) %>%
  summarise(cnt = n())


df_voter_hisotry_copy %>%
  filter(Election_Type %in% c("PP", "PR")) %>%
  select(county_emsid, gender, race_est, election_year) %>%
  distinct() %>%
  group_by(election_year) %>%
  summarise(cnt = n()) %>%
  ggplot(aes(election_year, cnt)) + 
  geom_line(color="grey") +
  geom_point(data=df_PP, shape=21, color="black", fill="#69b3a2", size=2) +
  theme_ipsum() +
  ggtitle("Number of voters voted in the primary elections each year") +
  labs(x = "year", y = "count")
  

```



